---
title: "Checking Primer Pairs"
output: html_notebook
---
Given what we know about whole COI, or whole mtgenomes , or whole 12S-16S sequences, do already developed primers amplify long mtochondrial fragments ? At least in silico


```{r}
library(rentrez)
library(insect)
library(tidyverse)
library(here)
library(gggenes)
```

# Download whole mitogenomes

```{r}
queries <- c("Polychaeta", "Vermetidae", "Gorgoniidae", "Pleurogona", 
             "Holothuriidae", "Bivalvia", "Asteroidea", 
             "Bryozoa", "Echinoidea", "Hexacorallia", "Scyllaridae")

new_queries2023 <- c("Tunicata",  "Palinuridae")

query <- paste0( new_queries2023, #queries,
                 '[ORGN]+AND+mitochondrion[TITL]+complete[TITL]')

Search3<- map(query, ~ searchGB(.x, taxIDs = T, prompt = F))

names(Search3) <- new_queries2023

Search3 %>% compact()-> goodies

 

for (i in 1:length(goodies)){
  writeFASTA(goodies[[i]], file = here("data", "Data_seq", "Claudio", paste0(names(goodies)[i], ".fasta")))
  
}
  map(~writeFASTA(.x, file = here("data", "Data_seq", "Moncho", "")))
```

```{r}
 Search3 %>%
  
  map(~tibble (names = names(.x)) ) %>% 
  set_names(nm = new_queries2023# queries
            ) %>% 
  bind_rows (.id = "query") %>% 
  distinct() %>% 
  separate(names, into = c('Acc', 'taxID'), sep = "\\|", convert = T) %>% 
  group_by(taxID) %>% 
  tally(name ="counts"  )
```


# Sets of primers

All primers are in 5->3, so we'll do linear searches instead of primer pairs

```{r}
seqs <-readFASTA(here("data", "all_align_withprimers_FyRrc.fas"))

names(seqs[14:31]) -> primers

primer_Seqs <- dna2char(seqs[14:31]) %>% str_remove_all("-")

names(primer_Seqs)<- primers

primer_Seqs <- c(primer_Seqs, LerayF = "GGWACWGGWTGAACWGTWTAYCCYCC",
                 "16Sar-L" = "CGCCTGTTTAACAAAAACAT", "LoboF" = "KBTCHACAAAYCAYAARGAYATHGG", 
                  "16SbrH_revcom" = "ACGTGATCTGAGTTCAGACCGG")

more_primers <- c( "M74_rev" = "TGATCTGAGTTCAAACCGGAG", "12Sai" ="AAACTAGGATTAGATACCCTATTAT" , "12Sbi_rev" = "ACACATCGCCCGTCGCTCTT")

cytb <- c("Cytb424F"= "GGWTAYGTWYTWCCWTGRGGWCARAT", "cobr825_rev"= "CAYATYAARCCRGARTGRTAYTT")

all_primers <- c(primer_Seqs, more_primers, cytb)

ordered_primers <- all_primers[c(selected_fwds, selected_revs)]

melting.temps <- read_csv2(here("data", "melting_temps.csv"))
```


```{r}
Search3 %>% # Start with the list of mitochondrial genomes
  compact() %>% 
  map(function(taxon){
    
    # all_primers %>% 
      ordered_primers %>% 
      map(~ virtualPCR(taxon, .x))
    
  } ) -> with.one.primer

write_rds(with.one.primer, here("data", "primer.search.rds"))
```

```{r}
with.one.primer <- read_rds(here("data", "primer.search.rds"))
```


<!-- ## Add the other primers -->

<!-- ```{r} -->
<!-- Search3 %>% # Start with the list of mitochondrial genomes -->
<!--   compact() %>%  -->
<!--   map(function(taxon){ -->

<!--     more_primers %>%  -->
<!--       map(~ virtualPCR(taxon, .x)) -->

<!--   } ) -> to.add -->

<!-- map2(with.one.primer, to.add, append) -> with.one.primer -->

<!-- ``` -->


<!-- ```{r} -->
<!-- Search3 %>% # Start with the list of mitochondrial genomes -->
<!--   compact() %>%  -->
<!--   map(function(taxon){ -->

<!--     cytb %>%  -->
<!--       map(~ virtualPCR(taxon, .x)) -->

<!--   } ) -> to.add -->

<!-- map2(with.one.primer, to.add, append) -> with.one.primer -->
<!-- ``` -->

### Summarising results

Let's try big bois: 


```{r}
# names(Search3) <- queries
# 
# Search3 %>% 
#   compact() %>% names() -> names(with.one.primer)

with.one.primer %>% 
  
   # map(compact) %>% 
  as_tibble %>% 
  mutate(loci = names(with.one.primer[[1]])) %>% 
  pivot_longer(-loci, names_to = "taxon") %>% 
  mutate(class = map_chr(value, class)) %>% 
  filter (class != "NULL") %>% 
  mutate(summary = map(value , ~tibble (names = names(.x), 
                                        sequence = dna2char(.x)) %>% 
                         mutate(length = str_length(sequence)))) %>% 
  select(loci, taxon, summary) %>% 
  unnest(summary) -> summary.data


write_csv(summary.data, here("data", "summary.data.csv"))

Search3 %>% 
  compact() %>% 
  map_df(~tibble(names = names(.x), 
                 sequence = dna2char(.x)) %>% 
        mutate(original.length = str_length(sequence)), .id = "taxon") %>% 
  ggplot (aes(x = original.length, fill = taxon)) + geom_histogram()+ facet_wrap(~taxon, scales = "free")
```


What I can do now is to use the length after and before trimming to know where each primer sits

## Length of sequences before trimming

```{r}




Search3 %>% 
  compact() %>% 
  map_df(~tibble(names = names(.x), 
                 sequence = dna2char(.x)) %>% 
        mutate(original.length = str_length(sequence)), .id = "taxon") %>% 
  select(-sequence) -> original.lengths


```

## Length of primers

```{r}
tibble(seqs = all_primers, loci = names(all_primers)) %>% 
  mutate(primer.length = str_length(seqs)) -> primer.lengths

primer.lengths %>% 
  mutate (gene = case_when(loci %in% c("16Sar-L", "16SbrH_revcom", "M74_rev") ~ "16S",
                           loci %in% c("12Sai", "12Sbi_rev")           ~ "12S",
                           loci %in% c("Cytb424F", "cobr825_rev")      ~ "Cytb",
                           TRUE                                    ~ "COI")) -> primer.lengths
```

## Combine lenghts to create final table

```{r}
summary.data %>% 
  select(loci, taxon, names, length) %>% 
  left_join(original.lengths) %>% 
  left_join(primer.lengths) %>% 
  select(-seqs) %>% 
  mutate(end  = original.length -length,
         start = end - primer.length) -> to.plot

to.plot %>% 
  group_by(names) %>% 
  filter(str_detect(gene, "COI")) %>%
  filter(start == min(start)) %>% 
  summarise(correction = 1490 - start) %>% 
  # ungroup() %>% 
  slice(1) %>% 
  left_join(.,to.plot) %>% 
  mutate(new.end = correction + end,
         new.start = correction + start,
         new.end = case_when(new.end < 0  ~ new.end + original.length,
                             TRUE                      ~ new.end),
         new.start = case_when(new.start < 0  ~ new.start + original.length,
                              TRUE                      ~ new.start)) -> to.plot

```

## Visualize only one entry per taxid
```{r}
to.plot %>% separate(names, into = c("Accesion", "taxid"), sep = "\\|") %>% 
  select(-Accesion) %>% 
  group_by(taxid, loci) %>% 
  nest() %>% 
  mutate(data = map(data, ~.x %>% slice(1))) %>% 
  unnest(data) -> plot
  
```

## Adding taxon info

You can get the taxonomy.rds file by using 'insect::taxonomy'


```{r}
# taxonomy <- insect::taxonomy()
# write_rds(taxonomy, here("data", "taxonomy.rds"))
taxonomy <- read_rds(here("data", "taxonomy.rds"))

plot %>% 
  ungroup() %>% 
  distinct(taxid) %>% 
  mutate(taxID = as.numeric(taxid)) %>% 
  select(-taxid) -> pre.lineages

insect::get_lineage(pre.lineages$taxID, db = taxonomy, cores = 'autodetect')-> lineages



extract_lineage <- function(taxid){
  
  taxid[c("class", "family", "genus")] %>% bind_rows()
}

lineages %>% 
  set_names(nm = pre.lineages$taxID) %>% 
  discard(function(x) is.na(x[[1]]) ) %>%summary.data
  map(extract_lineage) %>%
  bind_rows(.id = "taxid") %>% 
  select(taxid, class, family, genus) -> lineages_to_join
```


```{r}
left_join(plot, lineages_to_join) %>% 
  filter(taxon == "Polychaeta") %>% 
  ggplot(aes(xmin = new.start, xmax = new.end, y = taxid, fill = gene)) +
  geom_gene_arrow() +
  facet_wrap(~taxon, ncol = 1) +
  scale_fill_brewer(palette = "Set3") +
  geom_gene_label(aes(label = loci ), fontface = "bold") +
  theme_genes() +
  theme(axis.text.y = element_text(face = "italic"),
        axis.title.y = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(xlim = c(0, NA)) +
  labs(fill = "")

ggsave(filename = here("figures", "Poychaeta.png"),width = 24,dpi = "retina"
       )
```


## Align all primers

Reorder all COI primers so we can choose carefully


```{r}
to.plot %>% 
  filter (gene == "COI") %>%
  group_by(loci) %>% 
  summarise(med.start = median(new.start)) %>% 
  arrange(med.start) %>% 
  pull(loci) -> good.order.COI
```


```{r}
to.plot %>% separate(names, into = c("Accesion", "taxid"), sep = "\\|") %>% 
left_join(lineages_to_join) %>% 
  group_by(taxon) %>% 
  mutate (n_max_t = n_distinct(taxid),
          n_gen   = n_distinct(genus)) %>% 
  group_by(taxon, loci, gene) %>% 
  summarise(coverage = n_distinct(taxid)/n_max_t,
            genus    = n_distinct(genus)/n_gen,
            n_spp    = n_distinct(taxid),
            n_gen    = n_distinct(genus)) -> for.summary.plot

for.summary.plot %>% distinct() %>% 
  ggplot(aes(x= fct_relevel(loci,good.order.COI), y = taxon, fill = coverage)) +
  geom_tile(color = "black")+
  geom_label(aes(label = n_spp)) +
  scale_fill_viridis_c() +
  facet_grid(. ~ gene, scales = "free", space='free') +
  
  theme_dark() +
  theme(axis.text.x = element_text(angle = 45))

ggsave(filename = here("figures", "coverage_by_spp.png"), dpi = "retina", width = 24, height = 10)
```


## Distance between amplicons

```{r}
for.summary.plot %>%  
  
  mutate(dir = case_when(str_detect(loci, "rev") ~ "Reverse",
                         TRUE                    ~ "Forward")) %>% 
  group_by(taxon, gene, dir) %>% 
  arrange(desc(n_gen)) %>% 
  slice(1) %>% 
  left_join(primer.lengths) %>% 
  select(taxon, loci, gene, dir, seqs) %>% 
  pivot_wider(names_from = dir, values_from = c(seqs, loci)) -> ready.for.second


  

```

```{r}
to.plot %>% 
  group_by(names, correction, taxon, original.length) %>% 
  nest() %>% 
  mutate(data = map2(data,original.length,  ~.x %>% add_case(loci = "wholemt", gene = "wholemt", new.start = .y-100, new.end = .y))) %>% unnest(data) %>% 
  separate(names, into = c("Accession", "taxid"), sep = "\\|") %>% 
  
  group_by(taxon, loci, gene) %>% 
  summarise(new.end.median = median(new.end),
            new.start.median= median(new.start)) %>% 
  ggplot(aes(xmin = new.start.median, xmax = new.end.median+100, y = taxon, fill = gene)) +
  geom_gene_arrow(arrow_body_height = grid::unit(10, "mm")                  ) +
   # facet_wrap(~gene, nrow = 1) +
  scale_fill_brewer(palette = "Set3") +
  # geom_gene_label(aes(label = loci ), fontface = "bold") +
  theme_genes() +
  theme(axis.text.y = element_text(face = "italic"),
        axis.title.y = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(xlim = c(0, NA)) +
  labs(fill = "")

ggsave(here("figures", "average_loci_pos.png"),
       dpi = "retina",
       width = 28,
       height = 14)

```

Good effort, but we want to know, for each specimen, the distance between loci

```{r}
to.plot %>% 
  select(names, loci, taxon, original.length, new.start, gene) %>% 
  pivot_wider(names_from = c(gene, loci), values_from = new.start)
```

```{r}
to.plot %>% 
  select(names, loci, taxon, original.length, new.start, gene) %>% 
  group_by(names, gene) %>% 
  mutate(min = min(new.start),
         max = max(new.start)) %>% 
  # ungroup() %>% 
  filter (new.start == min | new.start == max) -> save.selected 
  
# save.selected %>% 
#   select(names, taxon, original.length, gene, min, max) %>% 
#   distinct() %>% 
#   filter(taxon == "Polychaeta") %>% 
#   ggplot(aes(xmin = min, xmax = max, y = names, fill = gene)) +
#   geom_gene_arrow(                  ) +
#    # facet_wrap(~gene, nrow = 1) +
#   scale_fill_brewer(palette = "Set3") +
#   # geom_gene_label(aes(label = gene ), fontface = "bold") +
#   theme_genes() +
#   theme(axis.text.y = element_text(face = "italic"),
#         axis.title.y = element_blank(),
#         legend.position = "bottom") +
#   coord_cartesian(xlim = c(0, NA)) +
#   labs(fill = "")


save.selected %>% 
  # select(names, taxon, original.length, gene, min, max) %>% 
  # distinct() %>% 
  filter (max -min > 5000)
pivot_wider(names_from = c(gene), values_from = c(min,max)) %>% 
  mutate(dist.COI.12S = max_12S - min_COI)
```
## Generate primers to order


```{r}
primer.lengths

selected_fwds <- c("FolmerF_LCO1490",
                   "LoboF",
                   "12Sai",
                   "Cytb424F",
                   "16Sar-L")

selected_revs <- c("FolmerR_HCO2198_revcomp",
                   "newPAT_R2_revcomp",
                   "12Sbi_rev",
                   "cobr825_rev",
                   "16SbrH_revcom")

cola_fwd <-"TCGTCGGCAGCGTCAGATGTGTATAAGAGACAG"
cola_rev <- "GTCTCGTGGGCTCGGAGATGTGTATAAGAGACAG"

fwds <- primer.lengths %>% 
  filter(loci %in% selected_fwds) %>% 
  mutate(with_tag = paste0(cola_fwd, seqs),
         name_tag = paste0(loci, "_illumina_F"))

seqinr::write.fasta(sequences = as.list(fwds$with_tag),
                    names = as.list(fwds$name_tag),
                    file.out = here("output", "fwds_tagged.fasta"))

revs <- primer.lengths %>% 
  filter(loci %in% selected_revs) %>% 
  mutate(original_or = insect::rc(seqs),
         with_tag = paste0(cola_rev, original_or),
         name_tag = str_replace(loci, "rev", "illumina_R"),
         name_tag = str_remove(name_tag, "comp|com"))


seqinr::write.fasta(sequences = as.list(revs$with_tag),
                    names = as.list(revs$name_tag),
                    file.out = here("output", "revs_tagged.fasta"))
```

## Selecion of candidates - performance together

```{r}
expand_grid(selected_fwds, selected_revs) %>% 
  left_join(fwds %>%  select(seqF = seqs, selected_fwds = loci)) %>% 
  left_join(revs %>% select(seqR = seqs, selected_revs = loci)) %>% 
  unite(seqF, seqR, sep = "...", col = "seq") %>% 
  unite(selected_fwds, selected_revs, sep = "-", col = "newname") -> for.cutadapt

for.cutadapt %>% 
  rowwise() %>% 
  mutate(files = map2(seq, newname, function(.x, .y){
    seqinr::write.fasta(sequences = .x, names = .y, file.out = here("output", "primers",paste0(.y, ".fasta")))
    
    
  }))


  
```



```{bash }

for file in ../data/Data_seq/Moncho/*fasta; do
basefile=$(basename "$file")
 for primer in ../output/primers/*fasta; do
 baseprimer=$(basename "$primer")
 cutadapt -g "file:$primer" -o "../output/after_pcrs/$basefile-$baseprimer" $file --discard-untrimmed
 done
 
 done


```



## Summarise the data

```{r}
all.files <- tibble(names = list.files(here("output", "after_pcrs")),
                    paths = list.files(here("output", "after_pcrs"), full.names = T))

safe_read <- possibly(readFASTA, otherwise = NA)
all.files %>% 
  separate(names, into = c("taxon", "rest"), remove = F, sep = ".fasta-") %>% 
  mutate(rest = str_replace(rest, "16Sar-L", "16SarL")) %>% 
  separate(rest, into = c("fwd", "rev"), sep = "-") %>% 
  mutate(rev = str_remove(rev, ".fasta")) %>% 
  mutate(seqs = map(paths, safe_read)) -> all.files

all.files %>% 
  filter(!is.na(seqs)) %>% 
  mutate(nseqs = map_dbl(seqs, length),
         length = map_df(seqs, function(.x) .x %>% map(length) %>% bind_rows() )) %>% 
  select(taxon, fwd, rev, length) %>% 
  unnest() %>% 
   pivot_longer(cols = where(is.numeric), names_to = "ACC", values_to = "len") %>% 
  filter (!is.na(len)) %>% 
  separate(ACC, into = c("Accesion", "taxid"), sep = "\\|") %>% 
left_join(lineages_to_join) %>% 
  group_by(taxon, fwd, rev) %>% 
  summarise(n_seqs = n(),
         n_genus = n_distinct(genus),
            avg.length =  mean (len, na.rm= T))  %>% 
  ggplot(aes(x = fwd, y = rev, fill= avg.length)) +
  geom_tile(color="black")+
  scale_fill_viridis_c() +
  geom_label(aes(label = round(avg.length,0)))+
  facet_wrap(~taxon)+
  theme_dark()

ggsave(filename = here("figures", "amplicon_length_by.taxa.png"), dpi = "retina", width = 24, height = 10)
```

