---
title: "Checking 12S db coverage"
output: html_notebook
---

In this project I am using the existing taxonomic classifier includding the FASTA sequences targetting 12S for vertabrate species (https://github.com/terrimporter/12SvertebrateClassifier/releases/tag/v2.0.0-ref). The aim is to compare this reference dataset with my species list from Cabañeros National Park to assess whether the species present in my study area are represented, providing a first estimate of coverage.

For full primer coverage I will test my primers "Mimammal" FW - GGGTTGGTAAATTTCGTGCCAGC RV -CATAGTGGGGTATCTAATCCCAGTTTG against the sequences

```{r}
library(tidyverse)
library(here)
library(eDNAfuns)
```

Target species

```{r}
species_df <- read_excel(here("data/Cabañeros lista de especies.xlsx")) %>%
  select(Genero_especie) %>%
  filter(!is.na(Genero_especie)) %>%
  mutate(Genero_especie = str_trim(Genero_especie))
```

Ref_db
```{r}
Vert_12S_db <- fasta_reader(here("dbs/dbs/v2.0.0-ref/mytrainseq.fasta")) %>% separate(header, into= c("Accession", "tax"), sep = "\t")
```
```{r}
Vert_12S_db <- Vert_12S_db %>% 
  mutate(species = str_extract(tax, "(?<=s__)[^;]+"),
         species = str_replace_all(species, "_", " "))
```
first check
```{r}
species_df %>% 
  inner_join(Vert_12S_db, by =c("Genero_especie"="species")) -> direct_matches

direct_matches %>% 
  distinct(Genero_especie)

species_df %>% 
  anti_join(Vert_12S_db, by =c("Genero_especie"="species")) %>% 
  separate(Genero_especie, into= c("genus", "species")) -> missing_taxa

Vert_12S_db %>% 
  mutate(genus = str_extract(tax, "(?<=g__)[^;]+")) -> Vert_12S_db

missing_taxa %>% 
   distinct(genus) %>% 
  inner_join(Vert_12S_db %>% select(genus, species)) -> with_genus_coverage

missing_taxa %>% 
   distinct(genus) %>%
  anti_join(Vert_12S_db )
```

Check rCRUX

```{r}
taxonomy_cols <- c("kingdom", "phylum", "class", "order", "family", "genus", "species")

rCRUX_tax <- read_delim(here("dbs/12S_expanded/derep_and_clean_db/12S_expanded_derep_and_clean_taxonomy.txt"), delim = "\t", quote = "\"", col_names = c("header" , "tax")) %>% separate(tax, into = taxonomy_cols, sep = ";")

rCRUX_tax %>% 
  distinct(species)
  
  
```
first check
```{r}
species_df %>% 
  inner_join(rCRUX_tax, by =c("Genero_especie"="species")) -> direct_matches

direct_matches %>% 
  distinct(Genero_especie)

missing_taxa <- species_df %>% 
  anti_join(rCRUX_tax, by = c("Genero_especie" = "species"))
```
Combinar ambas bases

```{r}
species_df_clean <- species_df %>% 
  mutate(species_join = tolower(trimws(Genero_especie)))

Vert_12S_db_clean <- Vert_12S_db %>% 
  mutate(species_join = str_extract(tax, "(?<=s__)[^;]+"),
         species_join = str_replace_all(species_join, "_", " "),
         species_join = tolower(trimws(species_join)))

rCRUX_tax <- rCRUX_tax %>% 
  mutate(species_join = tolower(trimws(species)))

ref_db <- bind_rows(
  Vert_12S_db_clean %>% select(species_join),
  rCRUX_tax %>% select(species_join)
) %>% distinct()

missing_taxa <- species_df_clean %>% 
  mutate (species_join= str_replace(species_join, "rana perezi", "pelophylax perezi")) %>% 
  anti_join(ref_db, by = "species_join")

```

FOr the missing ones, check GB to see if there are reference sequecnes that appear in GB but not yet in the curated databases

```{r}
target <- here("dbs/download_seqs/")

missing_taxa %>% 
  rowwise() %>% 
  mutate(query_12S = map_chr(Genero_especie, ~paste0(.x, "[ORGN]+AND+12S[TITL]"))) -> missing_taxa


search_safe <- function(q) {
  tryCatch(
    insect::searchGB(
      query = URLencode(q),
      taxIDs = TRUE,
      sequences = TRUE,
      prompt = FALSE
    ),
    error = function(e) NA,
    warning = function(w) NA
  )
}


missing_taxa %>% 
  mutate(seqs_12S = map(query_12S, ~search_safe(.x))) -> missing_taxa
```

```{r}
(missing_taxa$seqs_12S) %>% 
  map_chr(class) %>% 
  enframe() %>% 
  bind_cols(missing_taxa) %>% 
  
  filter(value != "DNAbin") -> still_missing
```

```{r}
(missing_taxa$seqs_12S) %>% 
  map_chr(class) %>% 
  enframe() %>% 
  bind_cols(missing_taxa) %>% 
  
  filter(value == "DNAbin") -> NCBI_for_gaps
```
Now let's see if the fragments contain our primers (it could be that the sequence is homologous but the submitters already deleted the primers)

```{r}
MiMammalF <- "GGGTTGGTAAATTTCGTGCCAGC"
MiMammalR <- "CATAGTGGGGTATCTAATCCCAGTTTG"
NCBI_for_gaps %>% 
  mutate (target = map(seqs_12S, ~virtualPCR(.x, up= MiMammalF)) )
```
keep sequences of the ones that were found - fasta folder 
modify the search for mitohchondrion genome - ve cambiando el query (que te diga moncho que ir poniendo)
base de satos de referencia y de las especies de datos que faltan tengo estas
de q especies hay cosas y de cuales no y explorar base de datos de mitogenomas en vez de 12S mitochondrion 
