---
title: "Gaps in database coveragek"
output: html_notebook
---
Search for species richness in a group, vs database coverage for major loci

```{r, message=F, echo=F}
library (tidyverse)
library (rentrez)
library(here)
library(insect)

```

## Mictophids

Mictophids are a very diverse group, but with patchy coverage in public databases

### Step1: species richness

Of course, the number of described species is not an accurate reflection of the actual species diversity in nature, but it will be the baseline for this experiment.

```{r}
taxonomy.today <- read_rds(file.path(Shared_resources,"taxonomy.rds" ))


```

### Get children taxa from Myctophids
```{r}
taxonomy.today %>% 
  filter (name == "Myctophidae") %>% 
  select(parent_taxID = taxID) %>% 
  semi_join(taxonomy.today, .) %>%# 32 genera
  select(parent_taxID = taxID, genus = name) %>%
  inner_join(taxonomy.today, .) -> Step1 # 233 species
  
Step1 %>% 
  filter (rank == "no rank") %>% 
  filter (!str_detect(name, "environmental")) %>% 
  select(parent_taxID = taxID, genus) %>%
  inner_join(taxonomy.today, .) %>% 
  filter (!str_detect(name, "environmental")) -> Bins # 131 bins

Step1 %>% 
filter (rank != "no rank") -> Myctophid.spp

bind_rows(Myctophid.spp, Bins) -> All.Myctophid

```

## Genebank Searches

Let's search the coverage of these species for the most common reference markers

### 16S rRNA

```{r}
query <- "Myctophidae[ORGN]+AND+16S[TITL]"
Mycto.GB <- searchGB(query,taxIDs = T, sequences = T, prompt = F )
# tibble(Accesions = Mycto.GB) ->
query2 <- "Myctophidae[ORGN]+AND+rRNA[TITL]"
Mycto2 <- searchGB(query2,taxIDs = T, sequences = T, prompt = F )
# tibble (Accessions = Mycto2)
query3 <- 'Myctophidae[ORGN]+AND+mitochondrion[TITL]'
Mycto3<- searchGB(query3, taxIDs = T, prompt = F)

```

#### Extract species info

```{r}
list (Mycto.GB, Mycto2, Mycto3) %>% 
  map(~tibble (names = names(.x)) ) %>% 
  bind_rows %>% 
  distinct() %>% 
  separate(names, into = c('Acc', 'taxID'), sep = "\\|", convert = T) %>% 
  group_by(taxID) %>% 
  tally(name ="counts"  ) -> Present.16S

Present.16S %>%
  right_join(All.Myctophid) %>% 
  replace_na(list (counts=0)) %>% 
  # distinct(counts)
  mutate (coverage = case_when(counts>1 ~ "+2" ,
                               TRUE     ~as.character(counts)),
          locus = "16S") %>% 
  select (name, genus, coverage, locus) %>% 
  mutate(coverage = fct_relevel(coverage, "+2", '1' )) -> to.plot
  # filter (genus %in% c("Benthosema", "Lobianchia", 'Diaphus')) %>% 
  
to.plot %>% 
  ggplot(aes(x = locus))+
  geom_bar(aes( fill = coverage)) +
  facet_grid(~genus, scales = "free", space = 'free' , )+
  coord_flip() +
  theme(panel.margin.x=unit(0, "lines"),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        ) +
  labs(x = '')

ggsave(filename = here("Manuscript", "16S.coverage.png"), height = 3, width = 21,dpi = 'retina')

to.plot %>%
  mutate(species = fct_reorder2(name, genus, coverage)) %>% 
  ggplot (aes(x = species)) +
  geom_bar(aes(fill = coverage), color = 'black')

ggsave(filename = here("Manuscript", "16S.coveragev2.png"), height = 3, width = 21,dpi = 'retina')
```

## Cytb rRNA

```{r}
query <- "Myctophidae[ORGN]+AND+Cytb[TITL]"
Mycto.GB <- searchGB(query,taxIDs = T, sequences = T, prompt = F )
```

```{r}
list (Mycto.GB, Mycto3) %>% 
  map(~tibble (names = names(.x)) ) %>% 
  bind_rows %>% 
  distinct() %>% 
  separate(names, into = c('Acc', 'taxID'), sep = "\\|", convert = T) %>% 
  group_by(taxID) %>% 
  tally(name ="counts"  ) -> Present.Cytb

Present.Cytb %>%
  right_join(All.Myctophid) %>% 
  replace_na(list (counts=0)) %>% 
  # distinct(counts)
  mutate (coverage = case_when(counts>1 ~ "+2" ,
                               TRUE     ~as.character(counts)),
          locus = "Cytb") %>% 
  select (name, genus, coverage, locus) %>% 
  mutate(coverage = fct_relevel(coverage, "+2", '1' )) -> to.plot
  # filter (genus %in% c("Benthosema", "Lobianchia", 'Diaphus')) %>% 
  
to.plot %>% 
  ggplot(aes(x = locus))+
  geom_bar(aes( fill = coverage)) +
  facet_grid(~genus, scales = "free", space = 'free' , )+
  coord_flip() +
  theme(panel.margin.x=unit(0, "lines"),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        ) +
  labs(x = '')

ggsave(filename = here("Manuscript", "Cytb.coverage.png"), height = 3, width = 21,dpi = 'retina')

to.plot %>%
  mutate(species = fct_reorder2(name, genus, coverage)) %>% 
  ggplot (aes(x = species)) +
  geom_bar(aes(fill = coverage), color = 'black')

ggsave(filename = here("Manuscript", "Cytb.coveragev2.png"), height = 3, width = 21,dpi = 'retina')
```

## 12S

```{r}
query <- "Myctophidae[ORGN]+AND+12S[TITL]"
Mycto.GB <- searchGB(query,taxIDs = T, sequences = T, prompt = F )
```
```{r}
list (Mycto.GB, Mycto2, Mycto3) %>% 
  map(~tibble (names = names(.x)) ) %>% 
  bind_rows %>% 
  distinct() %>% 
  separate(names, into = c('Acc', 'taxID'), sep = "\\|", convert = T) %>% 
  group_by(taxID) %>% 
  tally(name ="counts"  ) -> Present.12S

Present.12S %>%
  right_join(All.Myctophid) %>% 
  replace_na(list (counts=0)) %>% 
  # distinct(counts)
  mutate (coverage = case_when(counts>1 ~ "+2" ,
                               TRUE     ~as.character(counts)),
          locus = "12S") %>% 
  select (name, genus, coverage, locus) %>% 
  mutate(coverage = fct_relevel(coverage, "+2", '1' )) -> to.plot
  # filter (genus %in% c("Benthosema", "Lobianchia", 'Diaphus')) %>% 
  
to.plot %>% 
  ggplot(aes(x = locus))+
  geom_bar(aes( fill = coverage)) +
  facet_grid(~genus, scales = "free", space = 'free' , )+
  coord_flip() +
  theme(panel.margin.x=unit(0, "lines"),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        ) +
  labs(x = '')

ggsave(filename = here("Manuscript", "12S.coverage.png"), height = 3, width = 21,dpi = 'retina')

to.plot %>%
  mutate(species = fct_reorder2(name, genus, coverage)) %>% 
  ggplot (aes(x = species)) +
  geom_bar(aes(fill = coverage), color = 'black')

ggsave(filename = here("Manuscript", "12S.coveragev2.png"), height = 3, width = 21,dpi = 'retina')
```


## COI

```{r}
query <- "Myctophidae[ORGN]+AND+COI[TITL]"
Mycto.GB <- searchGB(query,taxIDs = T, sequences = T, prompt = F )
```

```{r}
list (Mycto.GB, Mycto3) %>% 
  map(~tibble (names = names(.x)) ) %>% 
  bind_rows %>% 
  distinct() %>% 
  separate(names, into = c('Acc', 'taxID'), sep = "\\|", convert = T) %>% 
  group_by(taxID) %>% 
  tally(name ="counts"  ) -> Present.COI

Present.Cytb %>%
  right_join(All.Myctophid) %>% 
  replace_na(list (counts=0)) %>% 
  # distinct(counts)
  mutate (coverage = case_when(counts>1 ~ "+2" ,
                               TRUE     ~as.character(counts)),
          locus = "COI") %>% 
  select (name, genus, coverage, locus) %>% 
  mutate(coverage = fct_relevel(coverage, "+2", '1' )) -> to.plot
  # filter (genus %in% c("Benthosema", "Lobianchia", 'Diaphus')) %>% 
  
to.plot %>% 
  ggplot(aes(x = locus))+
  geom_bar(aes( fill = coverage)) +
  facet_grid(~genus, scales = "free", space = 'free' , )+
  coord_flip() +
  theme(panel.margin.x=unit(0, "lines"),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        ) +
  labs(x = '')

ggsave(filename = here("Manuscript", "COI.coverage.png"), height = 3, width = 21,dpi = 'retina')

to.plot %>%
  mutate(species = fct_reorder2(name, genus, coverage)) %>% 
  ggplot (aes(x = species)) +
  geom_bar(aes(fill = coverage), color = 'black')

ggsave(filename = here("Manuscript", "COI.coveragev2.png"), height = 3, width = 21,dpi = 'retina')
```

```{r}
list ("12S" = Present.12S, "16S" = Present.16S, COI = Present.COI, Cytb = Present.Cytb) %>% 
  map(~right_join(.x,All.Myctophid)) %>% 
  bind_rows(.id = "locus") %>% 
  # right_join(All.Myctophid) %>% 
  replace_na(list (counts=0)) %>% 
  # distinct(counts)
  mutate (coverage = case_when(counts>1 ~ "+2" ,
                               TRUE     ~as.character(counts))) %>% 
  select (name, genus, coverage, locus) %>% 
  mutate(coverage = fct_relevel(coverage, "+2", '1' )) -> to.plot

write_rds(to.plot, here("Data", "coverage_data.rds"))
```

```{r}
to.plot <- read_rds( here("Data", "coverage_data.rds"))

to.plot %>% 
  ggplot(aes(x = locus))+
  geom_bar(aes( fill = coverage)) +
  facet_grid(~genus, scales = "free", space = 'free' , )+
  coord_flip() +
  theme(panel.margin.x=unit(0, "lines"),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        ) +
  labs(x = '')

ggsave(filename = here("Manuscript", "All.coverage.png"), height = 10, width = 21,dpi = 'retina')
```

```{r}
library(patchwork)

to.plot %>%
  group_by(locus) %>% 
  nest() %>% 
  mutate(data = map(data, ~ .x %>% 
                      mutate(species = fct_reorder2(name, genus, coverage))),
         plot = map2(data,locus ,  ~.x %>% 
                       mutate (locus = .y) %>% 
                       ggplot (aes(x = species)) +
  geom_bar(aes(fill = coverage), color = 'black') +
    theme(axis.title.x = element_blank()) +
  facet_wrap(ncol = 1, facets = ~locus)
                       ))   %>%
  pull(plot)-> plots 

plots[[4]] +   theme(axis.title.x = element_text(size = 18)) -> plots[[4]]

wrap_plots(plots,ncol = 1,guides = "collect") & 
  scale_fill_viridis_d(option = "plasma", direction = -1, breaks =c("0", "1", "+2") , labels = c("Absent", "1 sequence", "2 or more")) & 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        
        strip.text = element_text(face = "bold", color = "white", size = 22),
        strip.background = element_rect(fill = "black"),
        legend.position = "bottom",
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 24)) 
  
  

ggsave(filename = here("Manuscript", "All.coveragev2.png"), height = 11, width = 21,dpi = 'retina')
```

